---
title: "Assignment 5: Regression Part 2"
author: "Meghan Casey"
date: October 8, 2020
output: 
  html_document:
    theme: "cosmo" 
    toc: true
    toc_depth: 3
    toc_float: true
    highlight: "tango"
---

### Variables of Interest
Unit  of analysis: Household

Population: Renters in Massachusetts

1. Household Type (categorical): hh_type
2. Number of Units within Building (categorical): num_units
3. Age of Building (continuous): age_struct
4. Number of Bedrooms (continuous): bedrooms
5. Number of People (continuous): num_persons
6. Time Lived There (continuous): move_in
7. Gross Rent (continuous): gross_rent

I also added a new variable, density which is bedrooms/num_persons, so bedrooms per person.

### Load Libraries
```{r, message = FALSE}
library(tidyverse)
library(jtools)
library(interactions)
library(knitr)
```

### Loading Data
```{r, message = FALSE}
hh_data1 <- read_csv("households_ma.csv") %>%
   mutate(num_units = case_when(
    BLD_label=="2 Apartments"~"Fewer than 10 apartments",
    BLD_label=="10-19 Apartments"~"Ten to 49 apartments",
    BLD_label=="One-family house detached" ~ "1Single-family",
    BLD_label=="3-4 Apartments" ~ "Fewer than 10 apartments",
    BLD_label=="Mobile home or trailer" ~ "Mobile home or vehicle",
    BLD_label=="50 or more apartments" ~ "50+ apartments",
    BLD_label=="20-49 Apartments" ~ "Ten to 49 apartments",
    BLD_label=="5-9 Apartments"~"Fewer than 10 apartments",
    BLD_label=="One-family house attached" ~ "1Single-family",
    BLD_label=="Boat, RV, van, etc." ~ "Mobile home or vehicle")) %>%
  mutate(hh_type = case_when(
    HHT_label =="Married couple household"~"Married couple",
    HHT_label =="Other family household: Female householder, no spouse present"~"Single Parent",
    HHT_label =="Other family household: Male householder, no spouse present"~"Single Parent",
    HHT_label =="Nonfamily household: Male householder: Not living alone"~"NonFamily HH",
    HHT_label =="Nonfamily household: Female householder: Not living alone"~"NonFamily HH",
    HHT_label =="Nonfamily household: Female householder: Living alone"~"1Living Alone",
    HHT_label =="Nonfamily household: Male householder: Living alone"~"1Living Alone"))
```

#### Initial Regression
```{r}
model1 <- lm(gross_rent ~ age_struct + bedrooms + hh_type + num_units + move_in + num_persons, 
            data = hh_data1)

summary(model1)
```
### Additional Models
#### New Variable
I replaced num_persons and bedrooms with density.
```{r}
model2 <- lm(gross_rent ~ age_struct + density + hh_type + num_units + move_in, 
            data = hh_data1)

summary(model2)
```

#### Converting Continuous to Categorical
I've changed my age_struct variable to a categorical variable of New, Mid and Old aged. This increased the R-Squared from 0.22 to 0.23.
```{r}
hh_data2 <- hh_data1 %>%
  mutate(age_struct = case_when(age_struct <= 15 ~ "1New",
                                between(age_struct, 16, 79) ~ "Mid Age",
                                age_struct > 80 ~ "Old"))

model3 <- lm(gross_rent ~ age_struct + bedrooms + hh_type + num_units + move_in + num_persons, 
            data = hh_data2)

summary(model3)
```

#### Interaction Term
I used household type as an interaction term for density.

```{r}
hh_data3 <- hh_data1 %>%
  mutate(hh_type = case_when(
    HHT_label =="Married couple household"~"Family",
    HHT_label =="Other family household: Female householder, no spouse present"~"Family",
    HHT_label =="Other family household: Male householder, no spouse present"~"Family",
    HHT_label =="Nonfamily household: Male householder: Not living alone"~"NonFamily",
    HHT_label =="Nonfamily household: Female householder: Not living alone"~"NonFamily",
    HHT_label =="Nonfamily household: Female householder: Living alone"~"NonFamily",
    HHT_label =="Nonfamily household: Male householder: Living alone"~"NonFamily"))

model4 <- lm(gross_rent ~ age_struct + density + num_units + hh_type + (density:hh_type), 
            data = hh_data3)

summary(model4)
```

### Model Fit Visualizations
```{r}
ModelFit <- tibble(model = c(1, 2, 3, 4),
                   R_square = c(summary(model1)$adj.r.squared,
                                summary(model2)$adj.r.squared,
                                summary(model3)$adj.r.squared,
                                summary(model4)$adj.r.squared))

ggplot(ModelFit, aes(x = model, y = R_square)) +
  geom_line() +
  scale_x_continuous(name = "",
                   breaks = breaks <- seq(1, 4, by = 1),
                   labels = paste("Model", breaks)) +
  scale_y_continuous(name = "Adjusted R-squared value") +
  theme_bw()

```
```{r}

coeff_names <- c("Age of House" = "age_struct",
                 "Old House" = "age_structOld",
                 "Mid-Age House" = "age_structMid Age",
                 "50+ Units" = "num_units50+ apartments",
                 "<10 Units" = "num_unitsFewer than 10 apartments",
                 "Mobile Home or Vehicle" = "num_unitsMobile home or vehicle",
                 "10-49 Units" = "num_unitsTen to 49 apartments",
                 "Single Parent" = "hh_typeSingle Parent",
                 "Married Couple" = "hh_typeMarried couple", 
                 "NonFamily HH" = "hh_typeNonFamily HH",
                 "Number of Bedrooms" = "bedrooms",
                 "Number of People" = "num_persons",
                 "Time Lived There" = "move_in")

export_summs(model1, model3, 
             error_format = "(p = {p.value})",
             error_pos = "same",
             model.names = c("Initial model", "Preferred model"),
             coefs = coeff_names)
```
```{r}
coeff_desc <- c("Age of House" = "age_struct",
                 "Old House (compared to New)" = "age_structOld",
                 "Mid-Age House (compared to New)" = "age_structMid Age",
                 "50+ Units (compared to single family)" = "num_units50+ apartments",
                 "<10 Units (compared to single family)" = "num_unitsFewer than 10 apartments",
                 "Mobile Home or Vehicle (compared to single family)" = "num_unitsMobile home or vehicle",
                 "10-49 Units (compared to single family)" = "num_unitsTen to 49 apartments",
                 "Single Parent (compared to Living Alone)" = "hh_typeSingle Parent",
                 "Married Couple (compared to Living Alone)" = "hh_typeMarried couple", 
                 "NonFamily HH (compared to Living Alone)" = "hh_typeNonFamily HH",
                 "Number of Bedrooms" = "bedrooms",
                 "Number of People" = "num_persons",
                 "Time Lived There" = "move_in")

plot_coefs(model3, 
           coefs = coeff_desc,
           plot.distributions = TRUE,
           rescale.distributions = TRUE) +
  scale_x_continuous(name = "Effect on gross rent (dollars)")
```
```{r}
interact_plot(model4, pred = density, modx = hh_type, interval = TRUE) +
  scale_x_continuous(name = "Density of Household (Bedrooms per Person)",
                     limits = c(0, 6),
                     breaks = breaks <- seq(0, 6, by = 1)) +
  scale_y_continuous(name = "Gross Rent") +
  scale_color_discrete(name = "") +
  scale_linetype_discrete(name = "") +
  scale_fill_discrete(name = "") 

```