---
title: "Assignment 1: Identify a Dataset"
author: "Meghan Casey"
date: September 10, 2020
output: 
  html_document:
    theme: "cosmo" 
    highlight: "tango"
---
### Variables of Interest
Unit  of analysis: Household

Population: Renters in Massachusetts

1. Household Type (categorical): hh_type
2. Number of Units within Building (categorical): num_units
3. Age of Building (continuous): age_struct
4. Number of Bedrooms (continuous): bedrooms
5. Number of People (continuous): num_persons
6. Time Lived There (continuous): move_in
7. Gross Rent (continuous): gross_rent

### Load Libraries
```{r Load Libraries, message=FALSE} 
library(tidycensus)
library(tidyverse)
```

### Get Data
```{r Get Data, message=FALSE}
hh_vars_2018 <- pums_variables %>%
  distinct(year, survey, var_code, 
           var_label, data_type, level) %>%
  filter(level == "housing", year == 2018, survey == "acs1")
```

### Mass Renters Data Selection
```{r Mass Renters data filtered, message=FALSE, results='hide'} 
hh_data_MA <- get_pums(variables = c("TEN", 
                                  "YBL", 
                                  "BDSP", 
                                  "GRNTP", 
                                  "BLD",
                                  "NP",
                                  "MV",
                                  "HHT"),
                        state = "MA",
                        year = 2018, 
                        survey = "acs1",
                        recode = TRUE) %>%
  filter(TEN_label == "Rented") %>%
  mutate(age_struct = case_when(
    YBL_label == "1939 or earlier" ~ 100,
    YBL_label == "1940 to 1949" ~ 73,
    YBL_label == "1950 to 1959" ~ 63,
    YBL_label == "1960 to 1969" ~ 53,
    YBL_label == "1970 to 1979" ~ 43,
    YBL_label == "1980 to 1989" ~ 33,
    YBL_label == "1990 to 1999" ~ 23,
    YBL_label == "2000 to 2004" ~ 16,
    YBL_label == "2005" ~ 15,
    YBL_label == "2006" ~ 14,
    YBL_label == "2007" ~ 13,
    YBL_label == "2008" ~ 12,
    YBL_label == "2009" ~ 11,
    YBL_label == "2010" ~ 10,
    YBL_label == "2011" ~ 9,
    YBL_label == "2012" ~ 8,
    YBL_label == "2013" ~ 7,
    YBL_label == "2014" ~ 6,
    YBL_label == "2015" ~ 5,
    YBL_label == "2016" ~ 4,
    YBL_label == "2017" ~ 3,
    YBL_label == "2018" ~ 2)) %>%
  
  mutate(bedrooms = BDSP,
         gross_rent = GRNTP,
         num_persons = NP,
         hh_type = HHT,
         num_units = BLD)%>%
  
  mutate(move_in = case_when(
    MV_label == "12 months or less" ~ 1,
    MV_label == "13 to 23 months" ~ 2,
    MV_label == "2 to 4 years" ~ 3,
    MV_label == "5 to 9 years" ~ 7,
    MV_label == "10 to 19 years" ~ 15,
    MV_label == "20 to 29 years" ~ 25,
    MV_label == "30 years or more" ~ 30)) %>%
  select(age_struct, bedrooms, num_units, gross_rent, num_persons, move_in, hh_type)
```

I created my dataset that includes 17,922 observations, meaning 17,922 households. I looked at Sara Arman's rmd file to limit the preview of data to a specific number of rows. Here are the first 5 rows of data:
```{r}
head (hh_data_MA, 5)
```

### Save Dataset as CSV

```{r Save Dataset} 
write_csv(hh_data_MA, "households_ma.csv")
```
