---
title: "Assignment 2: Distributions and Proportions"
author: "Meghan Casey"
date: September 17, 2020
output: 
  html_document:
    theme: "cosmo" 
    highlight: "tango"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Variables of Interest
Unit  of analysis: Household

Population: Renters in Massachusetts

1. Household Type (categorical): hh_type
2. Number of Units within Building (categorical): num_units
3. Age of Building (continuous): age_struct
4. Number of Bedrooms (continuous): bedrooms
5. Number of People (continuous): num_persons
6. Time Lived There (continuous): move_in
7. Gross Rent (continuous): gross_rent

### Load Libraries
```{r Load Libraries, message = FALSE}
library(tidyverse)
library(ggplot2)
```

### Load Data
```{r Load Data, message = FALSE}
hh_data <- read_csv("households_ma.csv")
attach(hh_data)
```

## Distribution of Continuous Variables
### Summary of Statistics
```{r summary stats}
age_summary <- summary(age_struct)
age_sd <- sd(age_struct, na.rm = TRUE)
age_pop_mean_int <- t.test(age_struct)
bed_summary <- summary(bedrooms)
bed_sd <- sd(bedrooms, na.rm = TRUE)
bed_pop_mean_int <- t.test(bedrooms)
ppl_summary <- summary(num_persons)
ppl_sd <- sd(num_persons, na.rm = TRUE)
ppl_pop_mean_int <- t.test(num_persons)
live_summary <- summary(move_in)
live_sd <- sd(move_in, na.rm = TRUE)
live_pop_mean_int <- t.test(move_in)
rent_summary <- summary(gross_rent)
rent_sd <- sd(gross_rent, na.rm = TRUE)
rent_pop_mean_int <- t.test(gross_rent)
```

``` {r summary table}
summary_table <- tibble(
  Statistic = c("Sample mean", 
                "Median",
                "Standard deviation",
                "Interquartile range",
                "Population mean (95% confidence)"),
  `Age of Structure (years)` = c(prettyNum(age_summary["Mean"], digits = 3),
                           age_summary["Median"],
                           prettyNum(age_sd, digits = 3),
                           paste(age_summary["1st Qu."],"to", 
                                 age_summary["3rd Qu."]),
                           paste(prettyNum(age_pop_mean_int$conf.int[1],
                                           digits = 2), "to",
                                 prettyNum(age_pop_mean_int$conf.int[2],
                                               digits = 3))),
  
  `Number of Bedrooms` = c(prettyNum(bed_summary["Mean"], digits = 3),
                           bed_summary["Median"],
                           prettyNum(bed_sd, digits = 3),
                           paste(bed_summary["1st Qu."],"to", 
                                 bed_summary["3rd Qu."]),
                           paste(prettyNum(bed_pop_mean_int$conf.int[1],
                                           digits = 3), "to",
                                 prettyNum(bed_pop_mean_int$conf.int[2],
                                               digits = 3))),
  
  `Number of Residents` = c(prettyNum(ppl_summary["Mean"], digits = 3),
                           ppl_summary["Median"],
                           prettyNum(ppl_sd, digits = 3),
                           paste(ppl_summary["1st Qu."],"to", 
                                 ppl_summary["3rd Qu."]),
                           paste(prettyNum(ppl_pop_mean_int$conf.int[1],
                                           digits=3), "to",
                                 prettyNum(ppl_pop_mean_int$conf.int[2],
                                               digits=3))),
  
  `Years lived at House` = c(prettyNum(live_summary["Mean"], digits = 3),
                           live_summary["Median"],
                           prettyNum(live_sd, digits = 3),
                           paste(live_summary["1st Qu."],"to", 
                                 live_summary["3rd Qu."]),
                           paste(prettyNum(live_pop_mean_int$conf.int[1],
                                           digits = 3), "to",
                                 prettyNum(live_pop_mean_int$conf.int[2],
                                               digits = 3))),
  
  `Gross Rent` = c(prettyNum(rent_summary["Mean"], digits=3),
                           rent_summary["Median"],
                           prettyNum(rent_sd, digits = 3),
                           paste(rent_summary["1st Qu."],"to", 
                                 rent_summary["3rd Qu."]),
                           paste(prettyNum(rent_pop_mean_int$conf.int[1],
                                           digits=3), "to",
                                 prettyNum(rent_pop_mean_int$conf.int[2],
                                               digits=3))))

  knitr::kable(summary_table, 
             caption = "Characteristics of rented households in Massachusetts")
```

### Histograms

#### Age of Structure
```{r age_struct histogram}
ggplot(hh_data,
                   aes(x = age_struct)) +
  geom_histogram(color= "white", size = 1, fill = "paleturquoise4", binwidth = 5) +
  theme_gray() +
  scale_x_continuous(name = "Age of Building (years)") +
  scale_y_continuous(name = "Number of Houses in Sample") +
  annotate("text", x = 0, y = 6500, size = 3,
           label = paste("Population Mean (95% confidence): 63 to 63.9 years.
Sample Mean: 63.5 years.
Sample Median: 63 years.
Sample Standard Deviation: 32.3 years.
Sample Interquartile Range: 33 to 100 years."),
           hjust = 0, vjust = 1)

```


#### Number of Bedrooms
```{r bedrooms histogram}
ggplot(hh_data,
                   aes(x = bedrooms)) +
  geom_histogram(color= "white", size = 1, fill = "lightgoldenrod2", binwidth = 1) +
  theme_gray() +
  scale_x_continuous(name = "Number of Bedrooms in Household") +
  scale_y_continuous(name = "Number of Houses in Sample") +
  annotate("text", x = 3, y = 7000, size = 3,
           label = paste("Population Mean (95% confidence): 2.18 to 2.21 rooms.
Sample Mean: 2.2 rooms.
Sample Median: 2 rooms.
Sample Standard Deviation: 1.05 rooms.
Sample Interquartile Range: 2 to 3 rooms."),
           hjust = 0, vjust = 1)

```

The distribution of number of bedrooms is relatively normal, with a slight right skew. The mean is higher than the median due to the outliers at 6 bedrooms.

#### Number of People in Household
```{r people histogram}
ggplot(hh_data,
                   aes(x = num_persons)) +
  geom_histogram(color= "white", size = 1, fill = "azure3", binwidth = 1) +
  theme_gray() +
  scale_x_continuous(name = "Number of People in Household") +
  scale_y_continuous(name = "Number of Houses in Sample") +
  annotate("text", x = 7, y = 7000, size = 3,
           label = paste("Population Mean (95% confidence): 2.83 to 2.88 people.
Sample Mean: 2.85 people.
Sample Median: 3 people.
Sample Standard Deviation: 1.56 people.
Sample Interquartile Range: 2 to 4 people."),
           hjust = 0, vjust = 1)
```

The distribution of the number of people in households has a right skew with a long tail. Half of households consist of 2 to 4 people. 

#### Years lived at Household
```{r move histogram}
ggplot(hh_data,
                   aes(x = move_in)) +
  geom_histogram(color= "white", size = 1, fill = "honeydew3", binwidth = 3) +
  theme_gray() +
  scale_x_continuous(name = "Years Lived at Household") +
  scale_y_continuous(name = "Number of Houses in Sample") +
  annotate("text", x = 15, y = 7000, size = 3,
           label = paste("Population Mean (95% confidence): 5.6 to 5.79 years.
Sample Mean: 5.69 years.
Sample Median: 3 years.
Sample Standard Deviation: 6.54 years.
Sample Interquartile Range: 2 to 7 years."),
           hjust = 0, vjust = 1)
```

#### Gross Rent
```{r rent histogram}
ggplot(hh_data,
                   aes(x = gross_rent)) +
  geom_histogram(color= "white", size = 1, fill = "coral1", binwidth = 200) +
  theme_gray() +
  scale_x_continuous(name = "Gross Rent") +
  scale_y_continuous(name = "Number of Houses in Sample") +
  annotate("text", x = 15, y = 6500, size = 3,
           label = paste("Population Mean (95% confidence): $1,539 to $1,563 per month.
Sample Mean: $1,551 per month.
Sample Median: $1,440 per month.
Sample Standard Deviation: $825 per month.
Sample Interquartile Range: $980 to $2,020 per month."),
           hjust = 0, vjust = 1)
```

## Distribution of Categorical Variables
#### Household Types
```{r all possible values hh_type}
household_types <- unique(HHT_label)
household_types
```
```{r share hh_type}
married_couple <- t.test(HHT_label == household_types[5])
non_family_m <- t.test(HHT_label == household_types[1])
non_family_f <- t.test(HHT_label == household_types[2])
living_alone_m <- t.test(HHT_label == household_types[4])
living_alone_f <- t.test(HHT_label == household_types[2])
other_family_f <- t.test(HHT_label == household_types[6])
other_family_m <- t.test(HHT_label == household_types[7])
shares <- tibble(Household_Type = c("Married Couple",
                                    "Non-Family Male",
                                    "Non-Family Female",
                                    "Living Alone Male",
                                    "Living Alone Female",
                                    "Other Family Male",
                                    "Other Family Female"),
                 `Share` = c(married_couple$estimate,
                             non_family_m$estimate,
                             non_family_f$estimate,
                             living_alone_m$estimate,
                             living_alone_f$estimate,
                             other_family_m$estimate,
                             other_family_f$estimate),
                 Low = c(married_couple$conf.int[1],
                         non_family_m$conf.int[1],
                         non_family_f$conf.int[1],
                         living_alone_m$conf.int[1],
                         living_alone_f$conf.int[1],
                         other_family_m$conf.int[1],
                         other_family_f$conf.int[1]),
                  High = c(married_couple$conf.int[2],
                         non_family_m$conf.int[2],
                         non_family_f$conf.int[2],
                         living_alone_m$conf.int[2],
                         living_alone_f$conf.int[2],
                         other_family_m$conf.int[2],
                         other_family_f$conf.int[2]),)
knitr::kable(shares,caption = "Proportions and 95-Percent Confidence Intervals")
```

```{r bar chart hh_Type}
ggplot(shares, aes(x = household_types, y = Share, ymin = Low, ymax = High)) +
  geom_bar(stat = "identity", color = "white", size = 1, fill = "olivedrab4") +
  geom_errorbar(width = 0.5) +
  theme_gray() +
  scale_x_discrete(name = "Household Type", labels = c("Living Alone\n Female",
                                    "Non-Family\n Male",
                                    "Non-Family\n Female",
                                    "Living Alone\n Male",
                                    "Married Couple",
                                    "Other Family\n Male",
                                    "Other Family\n Female")) +
  scale_y_continuous(name = "Share of population (error bars represent\n 95-percent confidence interval)",
                     breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5),
                     labels = c("0", "10%", "20%", "30%", "40%", "50%"))
```

## Number of Units in Building
```{r all possible values units}
units <- unique(BLD_label)
units
```
```{r}
table(hh_data$BLD_label)
```

```{r share units}
two <- t.test(BLD_label == units[5])
three_4 <- t.test(BLD_label == units[2])
five_9 <- t.test(BLD_label == units[4])
ten_19 <- t.test(BLD_label == units[3])
twenty_49 <- t.test(BLD_label == units[6])
fifty <- t.test(BLD_label == units[1])
one_detached <- t.test(BLD_label == units[7])
one_attached <- t.test(BLD_label == units[8])
mobile <- t.test(BLD_label == units[9])
boat_rv <- t.test(BLD_label == units[10])
shares <- tibble(units = c("2 Apartments",
                                   "3-4 Apartments",
                                   "5-9 Apartments",
                                   "10-19 Apartments",
                                   "20-49 Apartments",
                                   "50 or more apartments",
                                   "One-family house detached",
                                   "One-family house attached",
                                   "Mobile home or trailer"   ,
                                   "Boat, RV, van, etc." ),
                 `Share` = c(two$estimate,
                             three_4$estimate,
                             five_9$estimate,
                             ten_19$estimate,
                             twenty_49$estimate,
                             fifty$estimate,
                             one_detached$estimate,
                             one_attached$estimate,
                             mobile$estimate,
                             boat_rv$estimate),
                 Low = c(two$conf.int[1],
                             three_4$conf.int[1],
                             five_9$conf.int[1],
                             ten_19$conf.int[1],
                             twenty_49$conf.int[1],
                             fifty$conf.int[1],
                             one_detached$conf.int[1],
                             one_attached$conf.int[1],
                             mobile$conf.int[1],
                             boat_rv$conf.int[1]),
                  High = c(two$conf.int[2],
                             three_4$conf.int[2],
                             five_9$conf.int[2],
                             ten_19$conf.int[2],
                             twenty_49$conf.int[2],
                             fifty$conf.int[2],
                             one_detached$conf.int[2],
                             one_attached$conf.int[2],
                             mobile$conf.int[2],
                             boat_rv$conf.int[2]),)
knitr::kable(shares,caption = "Proportions and 95-Percent Confidence Intervals")
```

```{r bar chart units}
ggplot(shares, aes(x = units, y = Share, ymin = Low, ymax = High)) +
  geom_bar(stat = "identity", color ="white", size = 1, fill = "tomato3") +
  geom_errorbar(width = 0.5) +
  theme_gray() +
  scale_x_discrete(name = "Number of Units in Building", labels = c("10-19\n Apts",
                                   "2\n Apts",
                                   "20-49\n Apts",
                                   "3-4\n Apts",
                                   "5-9\n Apts",
                                   "50 or more\n Apts",
                                   "Boat, RV,\n van, etc.",
                                   "Mobile\n or trailer",
                                   "One-fam\n attached",
                                   "One-fam\n detached")) +
  scale_y_continuous(name = "Share of population (error bars represent\n 95-percent confidence interval)",
                     breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5),
                     labels = c("0", "10%", "20%", "30%", "40%", "50%"))
```