---
title: "Assignment 4: Estimating a Regression Model"
author: "Meghan Casey"
date: October 1, 2020
output: 
  html_document:
    theme: "cosmo" 
    toc: true
    toc_depth: 4
    toc_float: true
    highlight: "tango"
---

### Variables of Interest
Unit  of analysis: Household

Population: Renters in Massachusetts

1. Household Type (categorical): hh_type
2. Number of Units within Building (categorical): num_units
3. Age of Building (continuous): age_struct
4. Number of Bedrooms (continuous): bedrooms
5. Number of People (continuous): num_persons
6. Time Lived There (continuous): move_in
7. Gross Rent (continuous): gross_rent

```{r message = FALSE}
library(tidyverse)
library(ggplot2)
library(ggthemes)
library(dbplyr)
library(dplyr)
```

```{r message = FALSE}
hh_data1 <- read_csv("households_ma.csv") %>%
   mutate(num_units = case_when(
    BLD_label=="2 Apartments"~"Fewer than\n10 apartments",
    BLD_label=="10-19 Apartments"~"10 to 49\napartments",
    BLD_label=="One-family house detached" ~ "Single-family",
    BLD_label=="3-4 Apartments" ~ "Fewer than\n10 apartments",
    BLD_label=="Mobile home or trailer" ~ "Mobile home\nor vehicle",
    BLD_label=="50 or more apartments" ~ "50+\napartments",
    BLD_label=="20-49 Apartments" ~ "10 to 49\napartments",
    BLD_label=="5-9 Apartments"~"Fewer than\n10 apartments",
    BLD_label=="One-family house attached" ~ "Single-family",
    BLD_label=="Boat, RV, van, etc." ~ "Mobile home\nor vehicle")) %>%
  mutate(hh_type = case_when(
    HHT_label =="Married couple household"~"1Married\ncouple",
    HHT_label =="Other family household: Female householder, no spouse present"~"Single\nParent",
    HHT_label =="Other family household: Male householder, no spouse present"~"Single\nParent",
    HHT_label =="Nonfamily household: Male householder: Not living alone"~"Roommates",
    HHT_label =="Nonfamily household: Female householder: Not living alone"~"Roommates",
    HHT_label =="Nonfamily household: Female householder: Living alone"~"Living\nAlone",
    HHT_label =="Nonfamily household: Male householder: Living alone"~"Living\nAlone"))
```

```{r}
model <- lm(gross_rent ~ age_struct + bedrooms + hh_type + num_persons, 
            data = hh_data1)

summary(model)
```