---
title: "Assignment 3: Correlations, Differences, + Associations"
author: "Meghan Casey"
date: September 24, 2020
output: 
  html_document:
    theme: "cosmo" 
    toc: true
    toc_depth: 4
    toc_float: true
    highlight: "tango"
---
### Variables of Interest
Unit  of analysis: Household

Population: Renters in Massachusetts

1. Household Type (categorical): hh_type
2. Number of Units within Building (categorical): num_units
3. Age of Building (continuous): age_struct
4. Number of Bedrooms (continuous): bedrooms
5. Number of People (continuous): num_persons
6. Time Lived There (continuous): move_in
7. Gross Rent (continuous): gross_rent

#### Load Libraries
```{r message = FALSE}
library(tidyverse)
library(ggplot2)
library(ggthemes)
```

#### Recode
```{r message = FALSE}
hh_data <- read_csv("households_MA.csv") %>%
   mutate(num_units = case_when(
    BLD_label=="2 Apartments"~"Building with fewer\nthan 10 apartments",
    BLD_label=="10-19 Apartments"~"Building with 10\nto 49 apartments",
    BLD_label=="One-family house detached" ~ "Single-family home",
    BLD_label=="3-4 Apartments" ~ "Building with fewer\nthan 10 apartments",
    BLD_label=="Mobile home or trailer" ~ "Mobile home or vehicle",
    BLD_label=="50 or more apartments" ~ "Building with 50+\napartments",
    BLD_label=="20-49 Apartments" ~ "Building with 10\nto 49 apartments",
    BLD_label=="5-9 Apartments"~"Building with fewer\nthan 10 apartments",
    BLD_label=="One-family house attached" ~ "Single-family home",
    BLD_label=="Boat, RV, van, etc." ~ "Mobile home or vehicle"))
```

### Relationships between two Continuous Variables
#### 1. Number of Bedrooms and Age of Structure
```{r}
correlation1 <- cor.test(hh_data$age_struct, hh_data$bedrooms)
correlation1

ggplot(hh_data, aes(x = bedrooms, y = age_struct)) +
  geom_point(size = 1, alpha = 0.3, position = "jitter", color = "tomato3") +
  theme_pander() +
  scale_y_continuous(name = "Age of Structure (years)") +
  scale_x_continuous(name = "Number of Bedrooms",
                     breaks = seq(0, 6, by = 1),
                     label = paste(prettyNum(seq(0, 6, by = 1)))) +
  
  annotate(geom = "text", x = 0, y = 90,
           label = paste("Correlation (95-percent confidence):",
                         prettyNum(correlation1$conf.int[1], digits = 2),
                         "to",
                         prettyNum(correlation1$conf.int[2], digits = 2)),
           hjust = 0)
```

#### 2.Number of People and Age of Building
```{r}
correlation2 <- cor.test(hh_data$age_struct, hh_data$num_persons)
correlation2

ggplot(hh_data, aes(x = num_persons, y = age_struct)) +
  geom_point(size = 1, alpha = 0.3, position = "jitter", color = "tomato3") +
  theme_pander() +
  scale_y_continuous(name = "Age of Structure (years)") +
  scale_x_continuous(name = "Number of People",
                     breaks = seq(0, 13, by = 1),
                     label = paste(prettyNum(seq(0, 13, by = 1)))) +
  
  annotate(geom = "text", x = 0, y = 90,
           label = paste("Correlation (95-percent confidence):",
                         prettyNum(correlation2$conf.int[1], digits = 2),
                         "to",
                         prettyNum(correlation2$conf.int[2], digits = 2)),
           hjust = 0)
```

#### 3. Time Lived There and Age of Building
```{r}
correlation3 <- cor.test(hh_data$age_struct, hh_data$move_in)
correlation3

ggplot(hh_data, aes(x = move_in, y = age_struct)) +
  geom_point(size = 1, alpha = 0.3, position = "jitter", color = "tomato3") +
  theme_pander() +
  scale_y_continuous(name = "Age of Structure (years)") +
  scale_x_continuous(name = "Time Lived There",
                     breaks = seq(0, 30, by = 5),
                     label = paste(prettyNum(seq(0, 30, by = 5)))) +
  
  annotate(geom = "text", x = 0, y = 90,
           label = paste("Correlation (95-percent confidence):",
                         prettyNum(correlation3$conf.int[1], digits = 2),
                         "to",
                         prettyNum(correlation3$conf.int[2], digits = 2)),
           hjust = 0)
```

#### 4. Gross Rent and Age of Building
```{r}
correlation4 <- cor.test(hh_data$age_struct, hh_data$gross_rent)
correlation4

ggplot(hh_data, aes(x = gross_rent, y = age_struct)) +
  geom_point(size = 1, alpha = 0.3, position = "jitter", color = "tomato3") +
  geom_smooth(color = "dimgray") +
  theme_pander() +
  scale_y_continuous(name = "Age of Structure (years)") +
  scale_x_continuous(name = "Gross Rent",
                     breaks = seq(0, 4500, by = 500),
                     label = paste("$", (prettyNum(seq(0, 4500, by = 500))))) +
  
  annotate(geom = "text", x = 0, y = 90,
           label = paste("Correlation (95-percent confidence):",
                         prettyNum(correlation4$conf.int[1], digits = 2),
                         "to",
                         prettyNum(correlation4$conf.int[2], digits = 2)),
           hjust = 0)

```

#### 5. Number of Bedrooms and Number of People
```{r}
correlation5 <- cor.test(hh_data$bedrooms, hh_data$num_persons)
correlation5

ggplot(hh_data, aes(x = bedrooms, y = num_persons)) +
  geom_point(size = 1, alpha = 0.3, position = "jitter", color = "tomato3") +
  theme_pander() +
  scale_y_continuous(name = "Number of People",
                     breaks = seq(0, 13, by = 2)) +
  scale_x_continuous(name = "Number of Bedrooms",
                     breaks = seq(0, 6, by = 1),
                     label = paste(prettyNum(seq(0, 6, by = 1)))) +
  
  annotate(geom = "text", x = -0.5, y = 13.5,
           label = paste("Correlation (95-percent confidence):",
                         prettyNum(correlation5$conf.int[1], digits = 2),
                         "to",
                         prettyNum(correlation5$conf.int[2], digits = 2)),
           hjust = 0)

```

#### 6. Number of Bedrooms and Time Lived There
```{r}
correlation6 <- cor.test(hh_data$bedrooms, hh_data$move_in)
correlation6

ggplot(hh_data, aes(x = bedrooms, y = move_in)) +
  geom_point(size = 1, alpha = 0.3, position = "jitter", color = "tomato3") +
  theme_pander() +
  scale_y_continuous(name = "Time Lived There",
                     breaks = seq(0, 35, by = 5),
                     label = paste(prettyNum(seq(0, 35, by = 5)))) +
  scale_x_continuous(name = "Number of Bedrooms",
                     breaks = seq(0, 6, by = 1),
                     label = paste(prettyNum(seq(0, 6, by = 1)))) +
  
  annotate(geom = "text", x = -0.5, y = 32,
           label = paste("Correlation (95-percent confidence):",
                         prettyNum(correlation6$conf.int[1], digits = 2),
                         "to",
                         prettyNum(correlation6$conf.int[2], digits = 2)),
           hjust = 0)
```

#### 7. Number of Bedrooms and Gross Rent
```{r}
correlation7 <- cor.test(hh_data$bedrooms, hh_data$gross_rent)
correlation7

ggplot(hh_data, aes(x = bedrooms, y = gross_rent)) +
  geom_point(size = 1, alpha = 0.3, position = "jitter", color = "tomato3") +
  theme_pander() +
  scale_y_continuous(name = "Gross Rent",
                     breaks = seq(0, 4500, by = 500),
                     label = paste("$", (prettyNum(seq(0, 4500, by = 500))))) +
  scale_x_continuous(name = "Number of Bedrooms",
                     breaks = seq(0, 6, by = 1),
                     label = paste(prettyNum(seq(0, 6, by = 1)))) +
  
  annotate(geom = "text", x = -0.5, y = 4600,
           label = paste("Correlation (95-percent confidence):",
                         prettyNum(correlation7$conf.int[1], digits = 2),
                         "to",
                         prettyNum(correlation7$conf.int[2], digits = 2)),
           hjust = 0)
```

#### 8. Number of People and Time Lived There
```{r}
correlation8 <- cor.test(hh_data$num_persons, hh_data$move_in)
correlation8

ggplot(hh_data, aes(x = move_in, y = num_persons)) +
  geom_point(size = 1, alpha = 0.3, position = "jitter", color = "tomato3") +
  theme_pander() +
  scale_y_continuous(name = "Number of People",
                     breaks = seq(1, 13, by = 2)) +
  scale_x_continuous(name = "Time Lived There",
                     breaks = seq(0, 35, by = 5),
                     label = paste(prettyNum(seq(0, 35, by = 5)))) +
  
  annotate(geom = "text", x = 13, y = 13.5,
           label = paste("Correlation (95-percent confidence):",
                         prettyNum(correlation8$conf.int[1], digits = 2),
                         "to",
                         prettyNum(correlation8$conf.int[2], digits = 2)),
           hjust = 0)
```

#### 9. Number of People and Gross Rent
```{r}
correlation9 <- cor.test(hh_data$num_persons, hh_data$gross_rent)
correlation9

ggplot(hh_data, aes(x = gross_rent, y = num_persons)) +
  geom_point(size = 1, alpha = 0.3, position = "jitter", color = "tomato3") +
  geom_smooth(color = "dimgray") +
  theme_pander() +
  scale_y_continuous(name = "Number of People",
                     breaks = seq(1, 13, by = 2)) +
  scale_x_continuous(name = "Gross Rent",
                     breaks = seq(0, 4500, by = 500),
                     label = paste("$", (prettyNum(seq(0, 4500, by = 500))))) +
  
  annotate(geom = "text", x = 2000, y = 13.5,
           label = paste("Correlation (95-percent confidence):",
                         prettyNum(correlation9$conf.int[1], digits = 2),
                         "to",
                         prettyNum(correlation9$conf.int[2], digits = 2)),
           hjust = 0)
```

#### 10. Time Lived There and Gross Rent
```{r}
correlation10 <- cor.test(hh_data$move_in, hh_data$gross_rent)
correlation10

ggplot(hh_data, aes(x = gross_rent, y = move_in)) +
  geom_point(size = 1, alpha = 0.3, position = "jitter", color = "tomato3") +
  geom_smooth(color = "dimgray") +
  theme_pander() +
  scale_y_continuous(name = "Time Lived There",
                     breaks = seq(0, 35, by = 5),
                     label = paste(prettyNum(seq(0, 35, by = 5)))) +
  scale_x_continuous(name = "Gross Rent",
                     breaks = seq(0, 4500, by = 500),
                     label = paste("$", (prettyNum(seq(0, 4500, by = 500))))) +
  
  annotate(geom = "text", x = 2000, y = 20,
           label = paste("Correlation (95-percent confidence):",
                         prettyNum(correlation10$conf.int[1], digits = 2),
                         "to",
                         prettyNum(correlation10$conf.int[2], digits = 2)),
           hjust = 0)
```

### Relationship between two Categorical Variables
#### Household Type and Number of Units
```{r}
chi_sq <- chisq.test(hh_data$HHT_label == "Married couple household", hh_data$num_units)
chi_sq

chi_sq$observed
chi_sq$expected
```

### Relationship between a continuous variable and a categorical variable (three or more levels)
#### 1. Age of Building and Household Type
```{r}
anova1 <- aov(age_struct ~ HHT_label, data = hh_data)
summary(anova1)

differences1 <- TukeyHSD(anova1)

as_tibble(cbind(pair = row.names(differences1$age_struct),
                differences1$HHT_label))

ggplot(hh_data, aes(x=HHT_label , y = age_struct)) +
  geom_boxplot() +
  theme_pander() +
  theme(axis.text.x = element_text(size = 10, angle = 45)) +
  scale_x_discrete(name = "Type of Household") +
  scale_y_continuous(name = "Age of Building (years)")
```

#### 2. Age of Building and Number of Units
```{r}
anova2 <- aov(age_struct ~ num_units, data = hh_data)
summary(anova2)

differences2 <- TukeyHSD(anova2)

as_tibble(cbind(pair = row.names(differences2$age_struct),
                differences2$num_units))

ggplot(hh_data, aes(y=num_units , x = age_struct)) +
  geom_boxplot() +
  theme_pander() +
  theme(axis.text.x = element_text(size = 10, angle = 45)) +
  scale_y_discrete(name = "Number of Units") +
  scale_x_continuous(name = "Age of Building (years)")
```

#### 3. Number of Bedrooms and Household Type
```{r}
anova3 <- aov(bedrooms ~ HHT_label, data = hh_data)
summary(anova3)

differences3 <- TukeyHSD(anova3)

as_tibble(cbind(pair = row.names(differences3$bedrooms),
                differences3$HHT_label))

ggplot(hh_data, aes(x=bedrooms , y = HHT_label)) +
  geom_boxplot() +
  theme_pander() +
  theme(axis.text.x = element_text(size = 12)) +
  scale_x_continuous(name = "Number of Bedrooms") +
  scale_y_discrete(name = "Household Type")
```

#### 4. Number of Bedrooms and Number of Units
```{r}
anova4 <- aov(bedrooms ~ num_units, data = hh_data)
summary(anova4)

differences4 <- TukeyHSD(anova4)

as_tibble(cbind(pair = row.names(differences4$bedrooms),
                differences4$num_units))

ggplot(hh_data, aes(x=bedrooms , y = num_units)) +
  geom_boxplot() +
  theme_pander() +
  theme(axis.text.x = element_text(size = 12)) +
  scale_x_continuous(name = "Number of Bedrooms") +
  scale_y_discrete(name = "Number of Units")
```

#### 5. Number of People and Household Type
```{r}
anova5 <- aov(num_persons ~ HHT_label, data = hh_data)
summary(anova5)

differences5 <- TukeyHSD(anova5)

as_tibble(cbind(pair = row.names(differences5$num_persons),
                differences5$HHT_label))

ggplot(hh_data, aes(x=num_persons , y = HHT_label)) +
  geom_boxplot() +
  theme_pander() +
  theme(axis.text.x = element_text(size = 12)) +
  scale_x_continuous(name = "Number of People") +
  scale_y_discrete(name = "Household Type")
```

#### 6. Number of People and Number of Units
```{r}
anova6 <- aov(num_persons ~ num_units, data = hh_data)
summary(anova6)

differences6 <- TukeyHSD(anova6)

as_tibble(cbind(pair = row.names(differences6$num_persons),
                differences6$num_units))

ggplot(hh_data, aes(x=num_persons , y = num_units)) +
  geom_boxplot() +
  theme_pander() +
  theme(axis.text.x = element_text(size = 12)) +
  scale_x_continuous(name = "Number of People") +
  scale_y_discrete(name = "Number of Units")
```

#### 7. Time Lived There and Household Type
```{r}
anova7 <- aov(move_in ~ HHT_label, data = hh_data)
summary(anova7)

differences7 <- TukeyHSD(anova7)

as_tibble(cbind(pair = row.names(differences5$move_in),
                differences5$HHT_label))

ggplot(hh_data, aes(x=move_in , y = HHT_label)) +
  geom_boxplot() +
  theme_pander() +
  theme(axis.text.x = element_text(size = 12)) +
  scale_x_continuous(name = "Time Lived There") +
  scale_y_discrete(name = "Household Type")
```

#### 8. Time Lived There and Number of Units
```{r}
anova8 <- aov(move_in ~ num_units, data = hh_data)
summary(anova8)

differences8 <- TukeyHSD(anova8)

as_tibble(cbind(pair = row.names(differences8$move_in),
                differences8$num_units))

ggplot(hh_data, aes(x=move_in , y = num_units)) +
  geom_boxplot() +
  theme_pander() +
  theme(axis.text.x = element_text(size = 12)) +
  scale_x_continuous(name = "Time Lived There") +
  scale_y_discrete(name = "Number of Units")
```
